R Notebook

### 2018-04-02 analysis/final_analysis.Rmd
Feature-based analyses

@@

```{r}
library(tidyverse)
library(ggthemes)
library(lme4)
library(lmerTest)

d_raw <- read_tsv('../data/final/experiment2/planned_sample/chatMessage/message.csv') %>%
  filter(iterationName == 'planned_sample') %>%
  filter(role == 'speaker')

# nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
incompleteIDs <- unique((d_raw %>% group_by(gameid) %>% 
                           filter(length(unique(trialNum)) != 24))$gameid)

# Some speakers relied purely on location in grid
pureLocation <- c('0180-a541e54c-5dea-4d06-9a21-9b842316eb59', '0101-0b659560-f05a-471a-89ac-60469e9c867f')
badGames <- c(incompleteIDs, pureLocation)

d <- d_raw %>%
  group_by(gameid, trialNum, occlusions, context) %>%
  summarize(text = paste0(text, collapse = ' ')) %>%   # Concatenate multiple utterance on same trial
  mutate(numRawWords = str_count(text, "\\S+")) %>%
  mutate(hidden = ifelse(occlusions == 'none', 'no', 'yes')) %>%
  filter(!(gameid %in% badGames))
```

@@

```{r}
basic.d <- d %>%
  filter(hidden == 'no') %>%
  group_by(gameid, context) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(context, m) %>%
  mutate(diff = close - far)

@@

```{r}
ggplot(basic.d, aes(x = diff)) +
    geom_histogram(binwidth = 0.25) +
    theme_bw() +
    geom_vline(xintercept = 0) +
    # ylim(1,12) +
    xlim(-4,4) +
    theme(aspect.ratio = 1)

t.test(basic.d$diff)
```

@@

d %>%
  group_by(context, hidden) %>%
  summarize(m = mean(numRawWords), se = sd(numRawWords)/sqrt(length(numRawWords))) %>%
  ggplot(aes(x = hidden, y = m, fill = context)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), position = dodge, width = 0.1) +
    theme_bw()
```

@@

```{r}
hidden.d <- d %>%
  group_by(gameid, context, hidden) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(hidden, m) %>%
  mutate(diff = yes - no)

@@

```{r}
ggplot(hidden.d %>% group_by(context) %>% mutate(m = mean(diff)), aes(x = diff)) +
    geom_histogram(binwidth = 0.33) +
    theme_bw() +
    geom_vline(xintercept = 0, linetype = 2) +
    geom_vline(aes(xintercept = m), color = 'red') +
    facet_wrap(~ context) +
    # ylim(1,12) +
    xlim(-5,5) +
    theme(aspect.ratio = 1)
```

@@

```{r}
summary(lmer(numRawWords ~ hidden + (1 + hidden | gameid), 
             data = d %>% 
                filter(context == 'far')))
t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far'), var.equal = FALSE)
```

@@

```{r}
summary(lmer(numRawWords ~ context * hidden + (1 + context * hidden | gameid), data = d))
```

@@

>>>>

## Import turk survey data

@@

```{r}
d_annotated <- read_csv('../data/final/experiment2/planned_sample/chatMessage/messages_with_annotations.csv')

# nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
incompleteIDs <- unique((d_raw %>% group_by(gameid) %>% 
                           filter(length(unique(trialNum)) != 24))$gameid)
confused <- unique((subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)

# Some speakers relied on location in grid (TODO: WHAT ABOUT 'BELOW' PEOPLE?)
location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
# Some speakers played this weird taboo game where they gave riddles 
tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
badGames <- c(incompleteIDs, pureLocation, confused)

d <- d_annotated %>%
  mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
  filter(!(gameid %in% badGames))
```

@@

```{r}
basic.d <- d %>%
  filter(hidden == 'no') %>%
  group_by(gameid, context) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(context, m) %>%
  mutate(diff = close- far,
         ratio = log(close / far))

@@

```{r}
ggplot(basic.d, aes(x = ratio)) +
    geom_histogram(binwidth = 0.25) +
    theme_bw() +
    geom_vline(xintercept = 0) +
    # ylim(1,12) +
    xlim(-4,4) +
    theme(aspect.ratio = 1)

t.test(basic.d$ratio)
```

@@

d %>%
  group_by(context, hidden) %>%
  summarize(m = mean(numRawWords), se = sd(numRawWords)/sqrt(length(numRawWords))) %>%
  ggplot(aes(x = context, y = m, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), position = dodge, width = 0.1) +
    theme_few() +
    ylab('mean # words') +
    theme(aspect.ratio = .5)

ggsave('~/Downloads/aggregated_means.pdf')
```

@@

Some people used exactly the same number of words on every trial...

```{r}
hidden.d %>% 
  filter(context == 'far') %>% 
  mutate(group = case_when(diff < -0 ~ 'anti',
                           diff > 0 ~ 'pro',
                           TRUE ~ 'all_the_things')) %>%
  group_by(group) %>% tally()

all_the_things_ers <- (hidden.d %>% filter(context =='far') %>% filter(diff == 0))$gameid
```

```{r}
hidden.d <- d %>%
  group_by(gameid, context, hidden) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(hidden, m) %>%
  mutate(diff = yes- no,
         ratio = log(yes/no))

@@

```{r}
ggplot(hidden.d %>% filter(!(gameid %in% all_the_things_ers)) %>% group_by(context) %>% mutate(m = mean(diff)), aes(x = diff)) +
    geom_histogram(binwidth = 0.33) +
    theme_bw() +
    geom_vline(xintercept = 0, linetype = 2) +
    geom_vline(aes(xintercept = m), color = 'red') +
    facet_wrap(~ context) +
    # ylim(1,12) +
    xlim(-6,6) +
    theme(aspect.ratio = 1)

ggsave("~/Downloads/histograms.pdf")
```

@@

```{r}
summary(lmer(numRawWords ~ hidden + (1 + hidden | gameid), 
             data = d %>% 
               filter(context == 'far') %>%
               filter(!(gameid %in% all_the_things_ers))
               ))
t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far') , var.equal = FALSE)
```

@@

```{r}
summary(lmer(numRawWords ~ context * hidden + (1 + context * hidden | gameid), data = d %>% filter(!(gameid %in% all_the_things_ers))))
```

@@

## Use feature data instead

```{r}
d %>% 
  group_by(context, occlusions) %>% 
  summarize(color = mean(colorMention), shape = mean(shapeMention), texture = mean(textureMention))
```

Histograms of number of features mentioned in each condition

```{r}
ggplot(d, aes(x = numFeatures)) +
  geom_histogram(binwidth = 1) +
  facet_grid(context ~ occlusions) +
  theme_few()
```

```{r}
d %>% group_by(context, occlusions) %>% 
  mutate(numTrials = length(text)) %>% 
  group_by(context, occlusions, colorMention, textureMention) %>% 
  summarize(n = length(text)/mean(numTrials)) %>%
  ggplot(aes(x = colorMention, y = textureMention, fill = n)) +
    geom_bin2d(stat = 'identity') +
    stat_bin2d(geom = "text", aes(label = round(n, 2))) +
    scale_fill_gradient(low = "white", high = "red") +
    facet_grid(occlusions ~ context) +
    theme_few()
```

Note: Random intercept doesn't converge (less variance in these numbers...)

```{r}
summary(lmer(numFeatures ~ context * occlusions + (1 + context + occlusions | gameid), data = d))
```

```{r}
summary(glmer(colorMention ~ occlusions + (1 + occlusions | gameid), family = 'binomial', data = d))
```

```{r}
summary(glmer(textureMention ~ occlusions + (1 + occlusions | gameid), family = 'binomial', data = d))
```

```{r}
d %>% 
  group_by(context, occlusions) %>% 
  summarize(numFeatures = mean(numFeatures))
```
